---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# DissimMapR

<!-- badges: start -->
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![test-coverage](https://github.com/macSands/DissimMapR/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/macSands/DissimMapR/actions/workflows/test-coverage.yaml)
<!-- badges: end -->

The goal of DissimMapR is to collate SUN deliverables for EU Horizon B3 project: Biodiversity Building Blocks for Policy

**Test git push**

## Installation

You can install the development version of DissimMapR from [GitHub](https://github.com/macSands/DissimMapR) with:

``` r
# install.packages("devtools")
devtools::install_github("macSands/DissimMapR")
```

## Example

This is a basic example which shows you how to simulate ecological data across a spatial and temporal grid, which can be used for testing spatial analysis methods for [b-cube](https://b-cubed.eu/):

```{r example}
library(DissimMapR)
```

#### Simulating Data
We will start by simulating a basic dataset with default parameters.

```{r}
# Set seed for reproducibility
set.seed(123)

# Simulate data
sim_data = dataSim(returnOption = 1, plot = FALSE)
head(sim_data)
```

#### Visualizing Data
If you want to visualize the spatial distribution of species counts, you can set the plot parameter to `TRUE`. Here we demonstrate this with `returnOption = 1`, which outputs a summary table of species by site and plots the data.

```{r plot-example, fig.width=7, fig.height=4}
# Generate data and plot
site_species = dataSim(returnOption = 1, plot = TRUE)
head(site_species)
```

#### Advanced Usage
The `dataSim()` function can also return a `SpatRaster` object or a distance matrix. Here's how to use these options:

```{r}
# Generate a SpatRaster object
raster_data = dataSim(returnOption = 2, plot = FALSE)
raster_data

# Calculate and return a distance matrix
distance_matrix = dataSim(returnOption = 3, plot = FALSE)
head(distance_matrix)
```

#### Saving Data to the Global Environment
If you need to access the raw grid data for further analysis, you can save it to the global environment using the globalSave option.

```{r}
site_species = dataSim(returnOption = 1, plot = FALSE, globalSave = TRUE)
# head(site_species)

# Accessing the grid data from global environment
head(dataSim_grid)
```

#### Conclusion
This vignette has introduced the basic functionalities of the `dataSim()` function. For more detailed information, please refer to the package documentation.
